---
title: "metaDataNew"
author: "Orfeas Gkourlias"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
library("plyr")
library(kableExtra)
library(ggplot2)
library(treemapify)
library(ggpubr)
library(stringr)
library(webshot)
library(magick)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)
rawDf <- read.delim("samplesWithPrediction_16_09_22_noOutliers.txt", sep = "\t", quote = "")
df <- rawDf
```

# Short Summary of selected RNA-Seq data sets.
The relevant file contains sample metadata originating from various tissues.

```{r}
colnames(df)[colnames(df) == "X"] <- "id"

kbl(df[0:5,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

### Tissue Prediction
Some of the tissue types had to be predicted. Most but not all predictions were
correct.

```{r echo=TRUE}
nrow(df[df$misclasified == "TRUE",])
df$Tissue <- df$predictedTissue
df[df$misclasified == "TRUE",]$Tissue <- df[df$misclasified == "TRUE",]$annotatedTissue
nrow(df)
```
Only 221 samples were incorrectly classified. The predicted tissues can simply be
replaced with the annotated ones.

### Tissue counts within studies.
To make sure the sample sizes for every tissue/project combination is enough, 30
was chosen as the threshold. This is how many samples will be omitted when
adhering to that threshold.



```{r echo=TRUE}
tissue_project <- df[,c("Tissue","study")]
tissue_per_project <- count(tissue_project)
tissue_per_project <- tissue_per_project[order(tissue_per_project$freq,decreasing=TRUE),]

tissues <- count(df[,c("Tissue")])
tissues <- tissues[order(tissues$freq,decreasing=TRUE),]
colnames(tissues)[colnames(tissues) == "x"] <- "Tissue"

tissues$Tissue <- str_replace(tissues$Tissue, "cardiac-organoids_cardiac progenitors_derived-cardiomyocytes", "Cardiomyocytes")
tissues$Tissue <- str_replace(tissues$Tissue, "Brain-Nucleus accumbens (basal ganglia)", "BNA basal ganglia")
tissues$Tissue <- str_replace(tissues$Tissue, "derived-neural-progenitor_derived-neurons", "derived-neurons")
tissues$Tissue <- str_replace(tissues$Tissue, "Salivary Gland-Minor Salivary Gland", "Salivary Gland Min.")
tissues$Tissue <- str_replace(tissues$Tissue, "fibroblasts_cell-lines_smooth-muscle-cell_mesenchymal-stem-cells", "cell_mesenchymal")
tissues$Tissue <- str_replace(tissues$Tissue, "derived-neural-progenitors_derived-astrocytes", "derived-astrocytes")
tissues$Tissue <- str_replace(tissues$Tissue, "cartilage/tenosynovium", "tenosynovium")
tissues$Tissue <- str_replace(tissues$Tissue, "Brain-Caudate (basal ganglia)", "BC basal ganglia")
tissues$Tissue <- str_replace(tissues$Tissue, "Brain-Cerebellar Hemisphere", "BCe. Hemis.")
tissues$Tissue <- str_replace(tissues$Tissue, "Brain-Spinal Cord (cervical c-1)", "Cerv. C-1")

tissues$Tissue <- str_wrap(tissues$Tissue,  whitespace_only = FALSE, width = 1)

projects <- count(df[,c("study")])
projects <- projects[order(projects$freq,decreasing=TRUE),]
colnames(projects)[colnames(projects) == "x"] <- "study"

cols <- colorRampPalette(brewer.pal(8, "Dark2"))(60)

ggplot(tissues, aes(area = freq, fill = Tissue,
               label = paste(Tissue, freq, sep = "\n"))) +
  geom_treemap() +
  geom_treemap_text(colour = "white",
                    place = "centre",
                    size = 24) +
  theme(legend.position = "none") +
  scale_fill_manual(values = cols)

ggsave(
  "plots/pbulic_distribution.png",
  plot = last_plot(),
  scale = 1,
  width = 24,
  height = 20,
  dpi = 300
)

# ggplot(tissues[1:27,], aes(x = Tissue, y = freq)) + 
#   geom_bar(stat = "identity", width = 0.75) +  
#   theme(axis.text.x = element_text(hjust=1)) +
#   coord_flip() +
#   labs(y = "Tissue frequency (n=46410)")
# 
# ggplot(tissues[28:49,], aes(x = Tissue, y = freq)) + 
#   geom_bar(stat = "identity", width = 0.75) +  
#   theme(axis.text.x = element_text(hjust=1)) +
#   coord_flip() +
#   labs(y = "Tissue frequency (n=46410)")
# 
# ggplot(tissues[50:60,], aes(x = Tissue, y = freq)) + 
#   geom_bar(stat = "identity", width = 0.75) +  
#   theme(axis.text.x = element_text(hjust=1)) +
#   coord_flip() +
#   labs(y = "Tissue frequency (n=46410)")

# kbl(tissues) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
#   scroll_box(height = "500px", width = "300px")
# 
# kbl(projects) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
#   scroll_box(height = "500px", width = "300px")
# 
# kbl(tissue_per_project) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
#   scroll_box(height = "500px", width = "500px")
```
Now for the study/tissue combinations with less than 30 samples.
```{r echo=TRUE}

left <- c(nrow(tissue_per_project[tissue_per_project$freq >= 60,]),
                nrow(tissue_per_project[tissue_per_project$freq >= 50,]),
                nrow(tissue_per_project[tissue_per_project$freq >= 40,]),
                nrow(tissue_per_project[tissue_per_project$freq >= 30,]),
                nrow(tissue_per_project[tissue_per_project$freq >= 20,]),
                nrow(tissue_per_project[tissue_per_project$freq >= 10,]))

lost <- nrow(tissue_per_project) - left

df$tissue_study <- paste(df$Tissue,df$study)
tissue_per_project$tissue_study <- paste(tissue_per_project$Tissue,tissue_per_project$study)

df_merge <- merge(df, tissue_per_project, by = "tissue_study")

samples_left <- c(nrow(df_merge[df_merge$freq >= 60,]),
                nrow(df_merge[df_merge$freq >= 50,]),
                nrow(df_merge[df_merge$freq >= 40,]),
                nrow(df_merge[df_merge$freq >= 30,]),
                nrow(df_merge[df_merge$freq >= 20,]),
                nrow(df_merge[df_merge$freq >= 10,]))

samples_lost <- nrow(df) - samples_left
study_left_df <- data.frame(threshold = c(60,50,40,30,20,10),
                            studies_left = left,
                            studies_lost = lost,
                            samples_left = samples_left,
                            samples_lost = samples_lost)

tissues_selected <- count(df_merge[df_merge$freq >= 30, c("Tissue")])
# kbl(tissue_per_project[tissue_per_project$freq < 30,]) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
#   scroll_box(height = "500px", width = "500px")
# 
# nrow(tissue_per_project[tissue_per_project$freq < 30,])
# nrow(tissue_per_project[tissue_per_project$freq >= 30,])
# kbl(tissue_per_project[tissue_per_project$freq < 30,]) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
#   scroll_box(height = "500px", width = "500px")

colnames(study_left_df) = c("Minimum Threshold", "Studies Discarded", "Studies Kept",
                            "Samples Discarded", "Samples Kept")

#study_left_df %>%
  #kbl(caption = "Threshold Summaries for Tissue/Study Frequencies") %>%
  #kable_classic(full_width = F, html_font = "Calibri") %>%
  #row_spec(seq(1, nrow(study_left_df), 1),extra_css = "border-bottom: 1px solid;") %>%
  #save_kable("plots/thresholds.pdf", density = 3000)


# kbl(study_left_df) %>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
#   save_kable("plots/thresholds.png")
```

That is 1797 Tissue/Study combinations with less than 30 samples.
There are 231 Tissue/Study combinations with 30 or more samples.
Which might look like a lot at first glance, but the ones with 30+
should contain many more samples.

```{r echo=TRUE}
sum(tissue_per_project$freq)
sum(tissue_per_project[tissue_per_project$freq < 30,]$freq)

sum(tissue_per_project[tissue_per_project$freq < 30,]$freq) / (sum(tissue_per_project[tissue_per_project$freq >= 30,]$freq) + sum(tissue_per_project[tissue_per_project$freq < 30,]$freq)) * 100
```
Combinations with more than or equal to 30 is 27222 while the
combinations with less than 30 samples covers 11007 samples, which is 28.79%`

Now to make the new dataframe without the lower sample counts.

We are now left with 27222 samples of which the study sizes are appropriate.

This is what the final data frame looks like.
```{r echo = TRUE}
z <- df
z[(z$study != "GTEx"),]$study <- "other"
y <- count(z[, c("Tissue", "study")])
y <- y[order(y$Tissue),]

nrow(z[z$study != "GTEx",])
nrow(z[z$study == "GTEx",])

tissue_count <- count(df$Tissue)
tissue_count <- tissue_count[order(tissue_count$freq, decreasing = TRUE),]

thresholds <- c(nrow(df[df$proj_tiss_freq < 60,]),
                nrow(df[df$proj_tiss_freq < 50,]),
                nrow(df[df$proj_tiss_freq < 40,]),
                nrow(df[df$proj_tiss_freq < 30,]),
                nrow(df[df$proj_tiss_freq < 20,]),
                nrow(df[df$proj_tiss_freq < 10,]))

percs <- round(thresholds/nrow(df) * 100, 0)


threshold_df <- data.frame(t(thresholds))
threshold_df[2,] <- t(percs)

colnames(tissue_count) = c("Tissue", "Frequency") 

colnames(threshold_df) <- c("threshold_60", "threshold_50", "threshold_40", "threshold_30", "threshold_20", "threshold_10")

rownames(threshold_df) <- c("n_samples", "percentage")

#ggplots
options(repr.plot.width = 5, repr.plot.height =2)

ggplot(y[1:27,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

ggplot(y[1:27,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

ggplot(y[28:49,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

ggplot(y[50:69,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

ggplot(y[70:89,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

kbl(tissue_per_project) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(height = "500px")


kbl(tissue_count) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(height = "500px")

kbl(threshold_df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(height = "500px")
```

Now to save the new dataframe to a file.
```{r}
write.csv(df, "selected_samples_final.csv")
tissue_project <- df[,c("Tissue","study")]
tissue_per_project <- count(tissue_project)
tissue_per_project <- tissue_per_project[order(tissue_per_project$freq,decreasing=TRUE),]
kbl(tissue_per_project[0:10,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(height = "500px")
```
```{r}

```

```{r}
three_proj_df <- df 
three_proj_df[(three_proj_df$study != "GTEx" & three_proj_df$study != "TCGA"),]$study <- "other"
gtex_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "GTEx",])
tcga_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "TCGA",])
other_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "other",])
gtex_count <- gtex_count[,c(1,3)]
tcga_count <- tcga_count[,c(1,3)]
other_count <- other_count[,c(1,3)]
colnames(gtex_count) <- c("Tissue", "gtex_freq")
colnames(tcga_count) <- c("Tissue", "tcga_freq")
colnames(other_count) <- c("Tissue", "other_freq")

df_list <- list(gtex_count, tcga_count, other_count)
merged_count <- Reduce(function(x, y) merge(x, y, all=TRUE), df_list, accumulate=FALSE)
merged_count
# 
merged_count[is.na(merged_count)] <- 0

merged_count$total <- merged_count$gtex_freq + merged_count$tcga_freq + merged_count$other_freq
kbl(merged_count) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(height = "500px")

write.csv(merged_count, "09-01-2024public_rna_tiss_proj_sum.csv")

y <- y[order(y$freq, decreasing = TRUE),]


ggplot(y[1:10,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

```
```{r}
merged_count <- merged_count %>%
  mutate(Tissue = ifelse(Tissue == "fibroblasts_cell-lines_smooth-muscle-cell_mesenchymal-stem-cells",
                         "Mesenchymal Cells",
                         Tissue))

merged_count <- merged_count %>%
  mutate(Tissue = ifelse(Tissue == "derived-neural-progenitor_derived-neurons",
                          "Derived Neurons",
                         Tissue))
merged_count <- merged_count[order(merged_count$total, decreasing = TRUE), ]
top_10 <- head(merged_count, 10)
top_10_long <- tidyr::pivot_longer(top_10, cols = c("gtex_freq", "tcga_freq", "other_freq"))

top_10_long$Tissue <- factor(top_10_long$Tissue, levels = top_10$Tissue)

colnames(top_10_long)[3] = "Study"

ggplot(top_10_long, aes(x = Tissue, y = value, fill = Study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Frequency") +
  ggtitle("Top 10 tissue frequencies (n=26,092)") +
  scale_fill_discrete(labels=c('GTEx', 'Other', 'TCGA'))

ggsave(
  "most.png",
  plot = last_plot(),
  scale = 1,
  width = 12,
  height = 8,
  dpi = 300
)
```
```{r}
studies <- projects[order(projects$freq, decreasing = TRUE),]
studies <- studies[1:10,]

transform(studies, study = reorder(study, freq))

ggplot(studies, aes(x =  reorder(study, freq), y = freq )) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Frequency") +
  ggtitle("Top 10 study  frequencies (n=16,045)")

ggsave(
  "studies.png",
  plot = last_plot(),
  scale = 1,
  width = 12,
  height = 8,
  dpi = 300
)
```
```{r}
df <- df[df$proj_tiss_freq >= 30,]
df
```


```{r}
three_proj_df <- df 
three_proj_df[(three_proj_df$study != "GTEx" & three_proj_df$study != "TCGA"),]$study <- "other"
gtex_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "GTEx",])
tcga_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "TCGA",])
other_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "other",])
gtex_count <- gtex_count[,c(1,3)]
tcga_count <- tcga_count[,c(1,3)]
other_count <- other_count[,c(1,3)]
colnames(gtex_count) <- c("Tissue", "gtex_freq")
colnames(tcga_count) <- c("Tissue", "tcga_freq")
colnames(other_count) <- c("Tissue", "other_freq")

df_list <- list(gtex_count, tcga_count, other_count)
merged_count <- Reduce(function(x, y) merge(x, y, all=TRUE), df_list, accumulate=FALSE)
merged_count
# 
merged_count[is.na(merged_count)] <- 0

merged_count$total <- merged_count$gtex_freq + merged_count$tcga_freq + merged_count$other_freq
kbl(merged_count) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  scroll_box(height = "500px")

write.csv(merged_count, "09-01-2024public_rna_tiss_proj_sum.csv")

y <- y[order(y$freq, decreasing = TRUE),]


ggplot(y[1:10,], aes(x = Tissue, y = freq, fill = study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Tissue frequency (n=46410)")

```
```{r}
three_proj_df <- df[df$proj_tiss_freq >= 30,] 
three_proj_df[(three_proj_df$study != "GTEx" & three_proj_df$study != "TCGA"),]$study <- "other"
gtex_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "GTEx",])
tcga_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "TCGA",])
other_count <- count(three_proj_df[,c("Tissue", "study")][three_proj_df$study == "other",])
gtex_count <- gtex_count[,c(1,3)]
tcga_count <- tcga_count[,c(1,3)]
other_count <- other_count[,c(1,3)]
colnames(gtex_count) <- c("Tissue", "gtex_freq")
colnames(tcga_count) <- c("Tissue", "tcga_freq")
colnames(other_count) <- c("Tissue", "other_freq")

df_list <- list(gtex_count, tcga_count, other_count)
merged_count <- Reduce(function(x, y) merge(x, y, all=TRUE), df_list, accumulate=FALSE)
merged_count
# 
merged_count[is.na(merged_count)] <- 0

merged_count <- merged_count %>%
  mutate(Tissue = ifelse(Tissue == "fibroblasts_cell-lines_smooth-muscle-cell_mesenchymal-stem-cells",
                         "Mesenchymal Cells",
                         Tissue))

merged_count <- merged_count %>%
  mutate(Tissue = ifelse(Tissue == "derived-neural-progenitor_derived-neurons",
                          "Derived Neurons",
                         Tissue))

merged_count$total <- merged_count$gtex_freq + merged_count$tcga_freq + merged_count$other_freq
merged_count
merged_count <- merged_count[order(merged_count$total, decreasing = TRUE), ]
top_10 <- head(merged_count, 10)
top_10_long <- tidyr::pivot_longer(top_10, cols = c("gtex_freq", "tcga_freq", "other_freq"))

top_10_long$Tissue <- factor(top_10_long$Tissue, levels = top_10$Tissue)

colnames(top_10_long)[3] = "Study"

ggplot(top_10_long, aes(x = Tissue, y = value, fill = Study)) + 
  geom_bar(stat = "identity", width = 0.75) +  
  theme(axis.text.x = element_text(hjust=1)) +
  coord_flip() +
  labs(y = "Frequency") +
  ggtitle("Top 10 tissue frequencies after threshold of 30 (n=19,850)") +
  scale_fill_discrete(labels=c('GTEx', 'Other', 'TCGA'))

ggsave(
  "most2.png",
  plot = last_plot(),
  scale = 1,
  width = 12,
  height = 8,
  dpi = 300
)
```
```{r}
colon_df <- df_merge[df_merge$Tissue.x == "Colon",]
colon_no_gtex <- colon_df[colon_df$study.x != "GTEx",]
colon_no_gtex_tcga <- colon_no_gtex[colon_no_gtex$study.x != "TCGA",]
write.csv(colon_no_gtex_tcga, "colon_pub_rna.csv")
```
